A Study of the Effectiveness of Usage Examples in REST API Documentation

### 2017-03-22 controlled_study.tex
remove authors for double-blind
fix-up

\ifCLASSINFOpdf
  \usepackage{listings}
  \usepackage[pdftex]{graphicx}
  \usepackage{url}
  % \usepackage{caption}
  \usepackage{subcaption}
  \usepackage{tikz,pgfplots,pgfplotstable}
  \usepackage{mathtools}
  \usepgfplotslibrary{statistics}
\else
\fi
\hyphenation{op-tical net-works semi-conduc-tor}

@@

\author{\IEEEauthorblockN{S M Sohan, Frank Maurer\IEEEauthorrefmark{1}}
\IEEEauthorblockA{\IEEEauthorrefmark{1}
Dept. of Computer Science\\
University of Calgary\\
Canada\\
\{smsohan, frank.maurer\}@ucalgary.ca}
\and
\IEEEauthorblockN{Craig Anslow}
\IEEEauthorblockA{School of Eng. and Computer Science\\
Victoria University of Wellington\\
New Zealand\\
craig@ecs.vuw.ac.nz}
\and
\IEEEauthorblockN{Martin Robillard}
\IEEEauthorblockA{
School of Computer Science\\
McGill University\\
Canada\\
martin@cs.mcgill.ca}
}

@@

\begin{abstract}
Generating and maintaining REST API documentation with usage examples can be a time consuming and expensive process for evolving APIs. Most REST API documentation tools focus on automating the documentation of the API objects, but require manual effort for capturing usage examples. Consequently, REST API developers need to know the cost vs. benefit of providing usage examples in the documentation to prioritize the documentation efforts. To this end, we have performed a controlled study with 26 experienced software engineers to understand problems that REST API client developers face while using an API without usage examples. We found that REST API client developers face productivity problems with using correct data types, data formats, required HTTP headers and request body when documentation lacks usage examples. By following the REST API documentation guidelines from this paper, REST API developers can reduce the errors, improve success rate and satisfaction of API client developers.

@@

\section{Research Question}
\label{sec:goal}
This research is aimed at answering the following research question:

@@

\section{Conclusion}
In this paper, we presented a set of problems that experienced REST API client developers face while performing API tasks using API documentation that lacks usage examples. We identified that, without examples, REST API client developers have trouble using the correct data types, correct data formats, and required HTTP headers and request body. We've also presented empirical evidence that by adding usage examples makes it possible to reduce mistakes, improve success rate and developer satisfaction of using the API. Practitioners can follow our implications while documenting REST APIs to improve API usability.

>>>>

\ifCLASSINFOpdf
  \usepackage{listings}
  \usepackage[pdftex]{graphicx}
  \usepackage{url}
  % \usepackage{caption}
  \usepackage{subcaption}
  \usepackage{tikz,pgfplots,pgfplotstable}
  \usepackage{mathtools}
  \usepgfplotslibrary{statistics}
\else
\fi

@@

\author{\IEEEauthorblockN{Placeholder Author Name}
\IEEEauthorblockA{
Placeholder Department\\
Placeholder Affiliation\\
Placeholder Address\\
Placeholder Email}
}

% \author{\IEEEauthorblockN{S M Sohan, Frank Maurer\IEEEauthorrefmark{1}}
% \IEEEauthorblockA{\IEEEauthorrefmark{1}
% Dept. of Computer Science\\
% University of Calgary\\
% Canada\\
% \{smsohan, frank.maurer\}@ucalgary.ca}
% \and
% \IEEEauthorblockN{Craig Anslow}
% \IEEEauthorblockA{School of Eng. and Computer Science\\
% Victoria University of Wellington\\
% New Zealand\\
% craig@ecs.vuw.ac.nz}
% \and
% \IEEEauthorblockN{Martin Robillard}
% \IEEEauthorblockA{
% School of Computer Science\\
% McGill University\\
% Canada\\
% martin@cs.mcgill.ca}
% }

@@

\begin{abstract}
Generating and maintaining REST API documentation with usage examples can be a time consuming and expensive process for evolving APIs. Most REST API documentation tools focus on automating the documentation of the API objects, but require manual effort for capturing usage examples. Consequently, REST API developers need to know the cost vs. benefit of providing usage examples in the documentation to prioritize the documentation efforts. To this end, we have performed a controlled study with 26 experienced software engineers to understand problems that REST API client developers face while using an API without usage examples. We found that REST API client developers face productivity problems with using correct data types, data formats, required HTTP headers and request body when documentation lacks usage examples. By following the REST API documentation suggestions from this paper, REST API developers can reduce the errors, improve success rate and satisfaction of API client developers.

@@

\section{Research Question}
\label{sec:goal}
This research is aimed at answering the following question:

@@

\section{Conclusion}
In this paper, we presented a set of problems that experienced REST API client developers face while performing API tasks using API documentation that lacks usage examples. We identified that, without examples, REST API client developers have trouble using the correct data types, correct data formats, and required HTTP headers and request body. We've also presented empirical evidence that by adding usage examples makes it possible to reduce mistakes, improve success rate and developer satisfaction of using the API. Practitioners can follow our implications as a set of guidelines while documenting REST APIs to improve API usability.
